---
title: "Data documentation"
author: "Leyla Nunez"
date: "1 april 2019"
output:
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

## Background

According to OP-507.01 - Study statistician has to add data specifications to the data documentation file (refer to TEMP-OP-507-A01 Data documentation). Data specifications are used to describe the contents of the data received form register holder including but not limited to: 

  - Name of data holder/source
  - Path to the raw data file in the server
  - Variable names in the data
  - Variable definitions
  - Definition of categories for factor variables
  - Coding system used (e.g. ICD-10, ICD-O-3, ATC, READ)
Variable structure (e.g. character, integer, date)

## Objective 

To avoit manual imputation to find an authomatic extraction of the information.

## Output

**Suggestion 1:** A **.txt** file will be saved. The layout used is the excel template **TEMP-OP-507-A01.01 Data documentation**. Why **.txt**? The file that needs to be exported is a list and I had some issues with succeeding to do so into **.xlsx**. 

**Suggestion 2:** An **.xlsx** file will be saved. The layout is changed, see the excel file. 

**Suggestion 3:** An **.html** file. The idea is that instead of creating excel sheets we could save the information required by the **OP-507.01** as a **html** file, alternatively as an **.pdf** file. The file can be copied to the project *../ER-xxx_Name_of_the_Study_DONOTTOUCH/Management/* folder in the Google drive. In addition, instead of having "Path to the raw" in the table, which makes the table long. I added that as a footnote to the table instead. I also suggest to remove the "comment". 
Now, I do thin that the information in the table should be limited. When we do data checks, we create an .R for each rawdata file. Those files could instead become RMarkdows. BUT I don't think this "data documentation" file should be as detailed as the individial data checks files. This one is just a "short" summary of the rawdata files delivered to us.


## Code

### Loading packages 
```{r load_packages, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(data.table)
library(knitr)
library(kableExtra)
```

### Adding some parameters
```{r parameters}

# The name of the rawdata folder. The default should be: 
# rawdata_folder_name <- "rawdata"
# I have 3 different test folders here
rawdata_folder_name <- "rawdata_test1"

# The exported txt/excel files will be saved here:
output_save_files <- here("report", "output", "data_check")

# the suggested column names for the output table
colnames <- c("Data holder/source", "Path to the raw data", "File",
              "Number of variables", "Number of observations",
              "Variable name", "Definition","Type", 
              "minimum", "maximum", "missing", "categories", "comments")

```


### Loading data

```{r loading_data}

# HERE: A list of all identified csv/txt files

# the absolute path
list_of_files_absPath <- list.files(path = here(rawdata_folder_name),
                            recursive = TRUE,
                            pattern = "\\.csv$|\\.txt$", 
                            full.names = TRUE) 

# the relative path
list_of_files_relPath <- list.files(path = here(rawdata_folder_name), 
                            recursive = TRUE,
                            pattern = "\\.csv$|\\.txt$", 
                            full.names = FALSE
                            ) 
# the relative path can also be extracted using:
# url_relative from xml2 package
library(xml2)
list_of_files_relPath <- url_relative(list_of_files_absPath, here(rawdata_folder_name, "."))

# Loading all the files and put them together into a LIST 
datalist <- lapply(list_of_files_absPath, function(x) fread(x,
                                                    header = TRUE,
                                                    stringsAsFactors = FALSE))


```

#### As a side point
In the code above I used **FRED** function to import *.csv*/*.txt* files. Additional extensions can of course be added, like Excel/SAS/SPSS/STATA/Rdata. For those extensions, other function for importing the files will be needed. In order to do so following steps can be created 


```{r as_a_side_point}

### 1:list ALL files 
all_files <- list.files(here(rawdata_folder_name), recursive = TRUE)

### 2: There are at least 2 ways to look for all the extrensions 
library(pathological)
unique(get_extension(list_of_files_relPath))
library(tools)
unique(file_ext(list_of_files_relPath))

### 3: Create a funtion that imports the files depending on the files extension

```

### Additional needed information

As I have seen, the rawdata files are usually located in subfolders to the **../rawdata**. In addition, the folder names are data holder/source names. In the code below, the data holder/source names will be extracted from the folder names. Now, if there are no subfolders the holder/source names in the outputs will be blank cells. 

```{r additional_info}
# Some additional information that will be used later on

### Name of data holder or source
dataHolder_source <- dirname(list_of_files_relPath) 
# If there is no subfolders in the ../rawdata/ folder the output will be "."
# > dataHolder_source
# [1] "." "." "." "." "." "."

# Names of each rawdata file - without the paths
name_rawdata <- basename(list_of_files_relPath) 
# The relative path to each of the rawdata files
path_rawdata <- here(rawdata_folder_name, dataHolder_source) # the relative path to the rawdata files

# beauty adjustment
dataHolder_source <- gsub("\\.", "Not specified", dataHolder_source)

```



### Suggestion 1

```{r data_extraction_v1, include=TRUE, echo=TRUE, warning=FALSE}

list_of_all_files <- NULL

for (i in seq_along(datalist)) { 

  # one df at the time  
  temp <- datalist[[i]]

  # extracting relevant information - more info can be added
  vari_type <- sapply(temp, class)
  vari_name <- names(temp)
  vari_def <- names(temp) #temporarily solution 
  
  vari_min <- sapply(temp, function(x) ifelse(is.numeric(x), min(x, na.rm = TRUE), NA))
  vari_max <- sapply(temp, function(x) ifelse(is.numeric(x), max(x, na.rm = TRUE), NA))
  vari_missing <- sapply(temp, function(x) sum(is.na(x)))
  vari_nrofcat <- sapply(temp, function(x) ifelse(is.character(x), length(unique(x)), NA))
  
  sanitary_check <- ifelse(is.na(vari_nrofcat), 
                           paste0("minimum: ", vari_min,
                                  "; maximum: ", vari_max,
                                  "; missing: ", vari_missing),
                           paste0("categories: ", vari_nrofcat,
                                  "; missing: ", vari_missing)
                           )
   
    file_info <- cbind(vari_name, 
             vari_def,
             vari_type,
             sanitary_check,
             comments = rep("", ncol(temp))
             )
    
    file_header <- rbind(c("", ""),
                        c(dataHolder_source[i], ""),
                        c(name_rawdata[i], path_rawdata[i]),
                        c(paste0("Number of variables: ", dim(temp)[2]), 
                          paste0("Number of observations: ", dim(temp)[1])),
                        c("", ""))

    list_of_all_files <- append(list_of_all_files,
                                list(file_header, file_info))

} 

# exporting to txt
lapply(list_of_all_files, function(x) write.table(data.frame(x),
                                                  file.path(output_save_files, "output_version1.txt"),
                                                  append = T, 
                                                  sep = '\t' , 
                                                  row.names = FALSE,
                                                  col.names = TRUE)
       )

```



### Suggestion 2

```{r data_extraction_v2, warning=FALSE}

list_of_all_files <- NULL


for (i in seq_along(datalist)) { 

  
    temp <- datalist[[i]]
  
  vari_type <- sapply(temp, class)
  vari_name <- names(temp)
  vari_def <- names(temp) #temporarily solution 
  
  minimum <- sapply(temp, function(x) ifelse(is.numeric(x), min(x, na.rm = TRUE), NA))
  maximum <- sapply(temp, function(x) ifelse(is.numeric(x), max(x, na.rm = TRUE), NA))
  missing <- sapply(temp, function(x) sum(is.na(x)))
  categories <- sapply(temp, function(x) ifelse(is.character(x), length(unique(x)), NA))

    file_info <- cbind(
             dataHolder_source = c(dataHolder_source[i], rep("", ncol(temp) - 1)),
             path_rawdata = c(path_rawdata[i], rep("", ncol(temp) - 1)),
             name_rawdata = c(name_rawdata[i], rep("", ncol(temp) - 1)),
             nr_rows = c(dim(temp)[1], rep("", ncol(temp) - 1)),
             nr_cols = c(dim(temp)[2], rep("", ncol(temp) - 1)),
             vari_name, 
             vari_def,
             vari_type,
             minimum,
             maximum,
             missing,
             categories,
             comments = rep("", ncol(temp))
             )
    
    list_of_all_files <- rbind(list_of_all_files, file_info)


} 

# Saving the file ...
write.table(list_of_all_files, 
            file.path(output_save_files, "output_version2.txt"),
            append = T, 
            sep = '\t' , 
            row.names = FALSE,
            col.names = colnames, 
            na = "")

```

### Suggestion 3

```{r creating table}
options(knitr.kable.NA = '') #Missing values (NA) in the table are displayed as NA by default. If you want to display them with other characters, you can set the option knitr.kable.NA, e.g. options(knitr.kable.NA = '') to hide NA values.

# removing "Path to the raw data" and "comments"" columns
list_of_all_files <- list_of_all_files[, -c(2, ncol(list_of_all_files))]
# adding column names
colnames(list_of_all_files) <- colnames[-c(2, length(colnames))]
# Adding an asterix to "Data holder/source" header
colnames(list_of_all_files)[1] <- paste0(colnames(list_of_all_files)[1],  
                                         footnote_marker_symbol(1))

list_of_all_files %>% 
  kable(row.names = FALSE,
        escape = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F, 
                position = "left",
              font_size = 10
              ) %>%
  #column_spec(2, width = "5cm",  bold = TRUE) %>% # was trying to contoll the width of "Path to the raw data" column. Now the column is removed and the paths are added as footnote to the table
  #collapse_rows(columns = 1:4, valign = "top") %>% # collapsing #Data holder/source" and "File" columns
  #scroll_box(width = "100%", height = "300px") %>% # does not work nicely with footnote
  footnote(general = "The path(s) to the raw data files.", 
           general_title = "General: ",
           number = paste0(unique(dataHolder_source), ": ", unique(path_rawdata)),
           footnote_as_chunk = F)

```


 